% -*- Mode: LaTeX; TeX-PDF-mode: t; -*-  # Config emacs auctex
\input{./._relpath-to-latexroot.ltx}    % Set paths to find resources (defines \latexroot and loads tex-paths.ltx)

\documentclass[titlepage, headings=optiontotocandhead]{econark} % embedded toc

% Definitions unique to this paper (showlabels config now in local.sty)
\usepackage{local}                % Local package definitions and setup
\usepackage{local-qe}             % local-qe gets pkgs req for qe

\bibliographystyle{plainnat}

% Load metadata BEFORE hypersetup (SST pattern)
\input{\latexroot/@local/metadata.ltx}

% When compiling Web version of paper, construct targets/anchors
\weborpdf{ % compiling web version
  \hypersetup{destlabel=true} % set up html labels
}{
  \provideboolean{showPageHead}{\setboolean{showPageHead}{true}}
  \hypersetup{
    pdftitle={\PDFTitle},
    pdfauthor={\PDFAuthor},
    pdfsubject={\PDFSubject},
    pdfkeywords={\PDFKeywords}
  }
}

\webonly{
  \section*{Endnotes}
  \theendnotes
}

\begin{document}

%% ----- Title page -----
\webonly{
  \newcommand{\versn}{Web}
}

\title{A Replication of Low (2005): Self-Insurance in a Life-Cycle Model of Labour Supply and Savings}

\author{REMARK Replication Package}

\keywords{Precautionary saving; Life-cycle labour supply}

\jelclass{D91; H31}

\date{}

\maketitle
\vspace{-0.5\baselineskip}

\whenintegrated{\label{abstract}} % integrated doc: SST for labels
\begin{abstract}
  \AbstractText
\end{abstract}
\vspace{-0.5\baselineskip}

This document is a REMARK-style replication note for Low (2005). Its purpose is to pair a
concise summary of the paper with executable code that reproduces the main quantitative
comparisons. In this repository, the computational replication is implemented in
\path{Code/Python/Low2005.py}.

\whenintegrated{\label{links}} % integrated doc: SST for labels

\providecommand{\REMARK}{\href{https://github.com/econ-ark/REMARK}{REMARK}}
\begin{footnotesize}
  \setlength{\parskip}{0pt}
  \parbox{0.9\textwidth}{
    \begin{tabbing}
      \texttt{~~~~~~~~~~~~~~~~~~~~~~~} \= \=  \\
      \texttt{~~~~GitHub~Repo:~~~~~~~} \> \> \texttt{\href{https://github.com/\owner/\projectname}{https://github.com/\owner/\projectname}} \\
      \texttt{~~~~~~~~~~~~~~~~~~PDF:~} \> \> \texttt{\href{https://github.com/\owner/\projectname/blob/main/build/Low2005.pdf}{build/Low2005.pdf}} \\
      \texttt{~~~~~~~~~~~~~~~~~~~~~~~} \> \>  \\
      \texttt{~~~~Replication~Package:~} \> \> \texttt{\href{https://doi.org/10.5281/zenodo.17861977}{10.5281/zenodo.17861977}} \\
      \texttt{~~~~~~~~~~~~~~~~~~~~~~~} \> \>  \\
      \texttt{~~~~Other~Materials:~~~} \> \> \\
      \texttt{~~~~~Replication~code:~} \> \> \texttt{\href{https://github.com/\owner/\projectname/tree/main/Code/Python}{Code/Python}} \\
    \end{tabbing}
  }
\end{footnotesize}
\vspace{-0.3\baselineskip}

\pagenumbering{gobble}

\titlepagefinish

\pagebreak

% Table of contents; to omit, \setboolean{TOC}{false} above
\ifthenelse{\boolean{TOC}}{\pagebreak
  \let\LaTeXStandardContentsName\contentsname
  \renewcommand{\contentsname}{}
  \tableofcontents
}{}

\section{Summary}
Low (2005) studies how households self-insure over the life cycle when future wages are
uncertain and markets are incomplete \cite{Low2005}. The paper's central point is that labour
supply is itself an insurance margin: if households can change hours worked, then they can
respond to wage risk using both savings and labour effort rather than savings alone.

This REMARK summarizes the paper briefly while leaving the main computational work to the
accompanying code. The Python implementation reproduces the paper's core comparisons
between certainty and uncertainty and between flexible and fixed labour supply.

The main economic conclusions are straightforward. Under uncertainty, households work more
and consume less early in life. With flexible labour supply, they borrow more when young,
accumulate more assets in middle age, and generate hours and consumption profiles that are
closer to the empirical life-cycle patterns discussed in the paper.

\section{Model Setup}
The model is a partial-equilibrium life-cycle problem in which households choose consumption,
leisure, and savings subject to wage risk. Period utility takes the Cobb--Douglas CRRA form
\begin{equation}
  u(c_t,l_t)=\frac{\left(c_t^{\eta} l_t^{1-\eta}\right)^{1-\gamma}}{1-\gamma}.
\end{equation}

Assets evolve according to
\begin{equation}
  A_{t+1}=(1+r)\left[A_t + (H-l_t)w_t - c_t\right],
\end{equation}
where $H-l_t$ is hours worked. The wage process combines a deterministic age profile with
permanent and transitory shocks:
\begin{align}
  \ln w_t &= \ln w_0 + \alpha_1 t + \alpha_2 t^2 + \ln w_t^P + \nu_t, \\
  \ln w_t^P &= \ln w_{t-1}^P + \varepsilon_t.
\end{align}

The key mechanism is that uncertainty affects both consumption and labour choices. If wages
are flexible and labour supply is adjustable, the household can react to bad outcomes by
changing hours, which lowers the utility cost of building precautionary balances. This is what
distinguishes the model from fixed-hours precautionary-savings frameworks such as
\cite{Deaton1991Low, GourinchasParker2002Low, Cagetti2003Low}.

\section{Calibration and Computation}
The paper calibrates the model at annual frequency and chooses the discount factor and the
utility weight on leisure to match average male hours and median wealth accumulation over
working life. The wage process moments are drawn from estimates in \cite{MeghirPistaferri2004Low},
and the key baseline values reported in the paper are summarized below.

\begin{table}[htbp]
  \centering
  \caption{Baseline calibration reported in Low (2005)}
  \begin{tabular}{lc}
    \toprule
    Parameter & Value \\
    \midrule
    Real interest rate $r$ & 0.016 \\
    Relative risk aversion $\gamma$ & 2.2 \\
    Consumption share $\eta$ & 0.4 \\
    Wage-growth coefficient $\alpha_1$ & 0.0561 \\
    Wage-growth coefficient $\alpha_2$ & -0.000599 \\
    Permanent-shock variance $\sigma^2_{\varepsilon}$ & 0.031 \\
    Transitory-shock variance $\sigma^2_{\nu}$ & 0.030 \\
    Social-security replacement rate & 0.55 \\
    \bottomrule
  \end{tabular}
\end{table}

The computational replication in this repository is implemented in
\path{Code/Python/Low2005.py}.
Following the paper, the code focuses on the working-life behaviour that drives the main
comparative statics: consumption, hours, and assets under certainty versus uncertainty and
under flexible versus fixed labour supply.

\section{Key Results}

\subsection{Flexible labour supply under certainty}
Under certainty, flexible hours generate a more pronounced life-cycle smoothing pattern than
fixed hours. Young households borrow more, middle-aged households save more, and hours
track the deterministic wage profile more closely. In the baseline utility specification,
consumption and leisure are Frisch substitutes, so the hump in hours worked induces a
corresponding hump in consumption.

\subsection{The effect of wage uncertainty}
Introducing uncertainty pushes labour supply upward early in life and lowers consumption at
young ages. The extra work effort creates a precautionary buffer of assets that can later be
used to smooth consumption and reduce hours when uncertainty has partially resolved. In the
paper, this mechanism helps explain why observed hours profiles are flatter than the wage
profile alone would suggest.

\subsection{Flexible versus fixed labour supply under uncertainty}
The key comparison in the paper is between economies with uncertain wages but different
degrees of labour-supply flexibility. Low (2005) finds that flexibility both changes the shape
of asset accumulation and raises the level of precautionary saving in the calibrated baseline.
Young households can postpone accumulation because they retain the option to adjust hours
later, while middle-aged households accumulate more assets before retirement.

\subsection{Alternative preferences}
The paper also checks robustness to alternative preferences. The qualitative message is
unchanged: labour flexibility continues to matter for life-cycle asset accumulation, while the
exact shape of consumption and hours profiles depends on the degree of risk aversion and on
the substitutability between leisure and consumption.

\begin{enumerate}
  \item Additive separability between consumption and leisure preserves the precautionary
  labour-supply motive, but changes the shape of consumption profiles.
  \item Varying $\gamma$ changes both the strength of precautionary motives and the degree of
  intertemporal substitution, with flatter hours profiles when risk aversion is higher.
  \item Replacing the Cobb--Douglas specification with a CES aggregator shows that the
  qualitative role of labour flexibility is robust, even though the magnitude of consumption and
  asset responses depends on the substitutability between leisure and consumption.
\end{enumerate}

\section{Conclusion}
Low (2005) shows that labour-supply flexibility is an important insurance margin
in life-cycle models with idiosyncratic wage risk. Ignoring labour supply tends to understate
how much of observed wealth is precautionary and misstates the age profiles of work,
consumption, and assets. In the calibrated simulations, flexibility leads households to borrow
more when young, save more in middle age, and produce life-cycle labour profiles that are
closer to the data. This is consistent with later quantitative work such as
\cite{French2004Low}, which also benefits from modelling labour supply jointly with savings
decisions.

\pagebreak

\clearpage
\section*{Appendices}
\addcontentsline{toc}{section}{Appendices}
\markboth{APPENDICES}{APPENDICES}

\appendix
\section{Appendix: Numerical Solution}
The original appendix solves the model recursively by backward induction over the life cycle.
After using the intratemporal first-order condition to express leisure as a function of
consumption and the wage, the dynamic problem can be reduced to solving for the
consumption rule on a finite grid of states. In the paper, expectations are evaluated
numerically and policy functions are interpolated between grid points. The HARK code in
this repository mirrors the economic comparisons of the paper while using modern solution
tools for the corresponding lifecycle household problem.

\end{document}
